---
title: "Computational Musicology"
author: "Gijs de Huu"
format: dashboard
---
# Structure Analysis
## Harmonic structure (Chroma Self-Similarity)
This chroma SSM displays patterns of harmonic repetition. Two of Hearts shows more harmonic cycles and Levo shows weaker harmonic repitition patterns which aligns with minimal techno’s focus on texture over harmonic progression.

```{r chroma-ssm-combined, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}

library(tidyverse)
library(compmus)

make_chroma_ssm <- function(path, label, downsample = 10L,
                            norm = "manhattan", dist = "cosine") {

  df <- read_csv(path, show_col_types = FALSE)

  chroma_cols <- c("C","C#","D","Eb","E","F","F#","G","Ab","A","Bb","B")

  dt <- median(diff(df$TIME), na.rm = TRUE)

  df %>%
    mutate(
      start = TIME,
      duration = lead(TIME) - TIME,
      duration = if_else(is.na(duration), dt, duration),
      chroma = pmap(select(., all_of(chroma_cols)), c)
    ) %>%
    select(start, duration, chroma) %>%
    filter(row_number() %% downsample == 0L) %>%
    mutate(chroma = map(chroma, compmus_normalise, norm)) %>%
    compmus_self_similarity(chroma, dist) %>%
    mutate(track = label)
}

two_chroma_ssm  <- make_chroma_ssm("two-chroma.csv",  "Two of Hearts", downsample = 10L)
levo_chroma_ssm <- make_chroma_ssm("levo-chroma.csv", "Levo",          downsample = 10L)

chroma_both <- bind_rows(two_chroma_ssm, levo_chroma_ssm)

p_chroma <- ggplot(chroma_both, aes(
  x = xstart + xduration/2,
  y = ystart + yduration/2,
  fill = d
)) +
  geom_tile(aes(width = 10 * xduration, height = 10 * yduration)) +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  facet_wrap(~ track, nrow = 1) +
  theme_void() +
  labs(title = "Harmonic Structure (Chroma Self-Similarity)", x = "", y = "")

p_chroma

dir.create("images", showWarnings = FALSE)

ggsave("images/chroma-ssm.png", p_chroma, width = 12, height = 5, dpi = 200)
```

## Timbral structure (MFCC)
These MFCC-based self-similarity matrices show a difference in structure. Where Levo shows clear changes in timbre, which is typical of minimal techno,
Two of Hearts appears more to be more similar in timbre.
```{r timbre-ssm-combined, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}

library(tidyverse)
library(compmus)

make_timbre_ssm <- function(path, label, downsample = 50L,
                            norm = "euclidean", dist = "cosine") {
  mfcc <- read_csv(path, show_col_types = FALSE)

  dt <- median(diff(mfcc$TIME), na.rm = TRUE)

  mfcc %>%
    mutate(
      start = TIME,
      duration = lead(TIME) - TIME,
      duration = if_else(is.na(duration), dt, duration),
      timbre = pmap(select(., starts_with("BIN")), c)
    ) %>%
    select(start, duration, timbre) %>%
    filter(row_number() %% downsample == 0L) %>%
    mutate(timbre = map(timbre, compmus_normalise, norm)) %>%
    compmus_self_similarity(timbre, dist) %>%
    mutate(track = label)
}

two_ssm  <- make_timbre_ssm("two-mfcc.csv",  "Two of Hearts", downsample = 50L)
levo_ssm <- make_timbre_ssm("levo-mfcc.csv", "Levo",          downsample = 50L)

ssm_both <- bind_rows(two_ssm, levo_ssm)

p_timbre <- ggplot(ssm_both, aes(
  x = xstart + xduration/2,
  y = ystart + yduration/2,
  fill = d
)) +
  geom_tile(aes(width = 50 * xduration, height = 50 * yduration)) +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  facet_wrap(~ track, nrow = 1) +
  theme_void() +
  labs(title = "Timbre Structure (MFCC Self Similarity)", x = "", y = "")

p_timbre
ggsave("images/timbre-ssm.png", p_timbre, width = 12, height = 5, dpi = 200)
```

# Harmonic Analysis
## Key Estimation
Two of Hearts shows clearer clustering around specific tonal centres and Levo shows a less clearly defined key pattern. 
```{r key-estimation, message=FALSE, warning=FALSE}
circshift <- function(v, n) {
  if (n == 0) v else c(tail(v, n), head(v, -n))
}

major_key <- c(6.35, 2.23, 3.48, 2.33, 4.38, 4.09, 2.52, 5.19, 2.39, 3.66, 2.29, 2.88)
minor_key <- c(6.33, 2.68, 3.52, 5.38, 2.60, 3.53, 2.54, 4.75, 3.98, 2.69, 3.34, 3.17)

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )

make_keygram <- function(path, label, every_n = 50L) {
  read_csv(path, show_col_types = FALSE) |>
    compmus_wrangle_chroma() |>
    filter(row_number() %% every_n == 0L) |>
    compmus_match_pitch_template(
      key_templates,
      method = "euclidean",
      norm = "manhattan"
    ) |>
    mutate(track = label)
}

two_keys  <- make_keygram("two-chroma.csv", "Two of Hearts", every_n = 50L)
levo_keys <- make_keygram("levo-chroma.csv", "Levo", every_n = 50L)

key_both <- bind_rows(two_keys, levo_keys)

p_key <- ggplot(
  key_both,
  aes(x = start + duration / 2, width = 50 * duration, y = name, fill = d)
) +
  geom_tile() +
  scale_fill_viridis_c(guide = "none") +
  facet_wrap(~ track, nrow = 1) +
  theme_minimal() +
  labs(title = "Estimated Key Over Time", x = "Time (s)", y = "")
```
